### 2015-12-11 conclusions.tex
Reword and reorganize some sentences

We have investigated scale-free algorithms for online linear optimization and
we have shown that scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors, i.e. without any assumption on the norms of the loss vectors
and the number of rounds. We have designed two scale-free algorithms for online linear
optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T
\|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly convex
function with respect to a norm $\|\cdot\|$ defined on the decision set and
where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
based on \textsc{Follow The Regularizer Leader}.  We have also designed an
algorithm with regret $O\left((R(u) + \frac{1}{\lambda}) \sqrt{T}
\max_{t=1,2,\dots,T} \|\ell_t\|_* \right)$ with respect to any competitor $u
\in K$, where $R$ is any non-negative $\lambda$-strongly convex function
defined on the decision set.  The latter result makes sense even when the
decision set $K$ is unbounded.

Similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
we gave examples of regularizers for which the associated Bregman divergence is
unounded and the scale-fre version of \textsc{Mirror Descent} has $\Omega(T)$
regret or worse.

We have shown an $\Omega\left( D \sqrt{\sum_{t=1}^T \|\ell\|_*^2}\right)$
lower bound on regret of any algorithm for decision set with diameter $D$.

>>>>

We have investigated scale-free algorithms for online linear optimization and
we have shown that scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors, i.e. without any assumption on the norms of the loss vectors
and the number of rounds. We have designed two scale-free algorithms for online
linear optimization with regret bound $O \left(\sqrt{\sup_{v \in K} f(v)
\sum_{t=1}^T \|\ell\|_*^2} \right)$ where $f$ is any non-negative $1$-strongly
convex function with respect to a norm $\|\cdot\|$ defined on the decision set
and where $\|\cdot\|_*$ is the dual norm to $\|\cdot\|$.  These algorithms are
based on \textsc{Follow The Regularizer Leader}.  We have shown that one of the
algorithms has regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} +
\max_{t=1,2,\dots,T} \|\ell_t\|_* \sqrt{T} \right)$ with respect to any
competitor $u \in K$, where $f$ is any non-negative $1$-strongly convex
function defined on the decision set.  The latter result makes sense even when
the decision set $K$ is unbounded.

Similar, but weaker result holds for a scale-free algorithm based on
\textsc{Mirror Descent}. However, we have also shown this algorithm is strictly
weaker than algorithms based on \textsc{Follow The Regularizer Leader}. Namely,
we gave examples of regularizers for which the associated Bregman divergence is
unbounded and the scale-free version of \textsc{Mirror Descent} has $\Omega(T)$
regret or worse.

We have shown an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
lower bound on regret of any algorithm for decision set with diameter $D$.

### 2015-12-11 introduction.tex
Reword and reorganize some sentences

Majority of existing algorithms for OLO are based on two generic algorithms:
\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
developed by \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed by~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and later
rediscovered in~\cite{Xiao-2010} for online convex optimization. The general
version with time-varying regularizers was analyzed in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
for convex optimization. In the online literature it makes the first
appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. Good
description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

>>>>

Majority of existing algorithms for OLO are based on two generic algorithms:
\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
forecaster in \cite[Chapters~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and later
rediscovered in~\cite{Xiao-2010} for online convex optimization. The general
version with time-varying regularizers was analyzed in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
for convex optimization. In the online learning literature it makes the first
appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. Good
description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

### 2015-12-11 mirror-descent.tex
Reword and reorganize some sentences

\section{Mirror Descent}
\label{section:mirror-descent}

@@

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The Regularized Leader} but
quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the Bregman
divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

@@

Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
prominent example is \textsc{Projected Gradient Descent} defined by the recurrence
$$
w_{t+1} = \Pi_K(w_t - \eta_t \ell_t)
$$
where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
an instance of \textsc{Mirror Descent} with regularizer $R_t(w) =
\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of projected gradient descent with
step size $\eta_t = \frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized
Infinitesimal Gradient Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}.
Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
\textsc{Mirror Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. When $d=1$, both algorithms are equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
generalization of Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2
= 1$ for all $i$, \textsc{Mirror Descent} with regularizer
\eqref{equation:ada-grad-regularizer} coincides with GIGA.

\textsc{Exponentially-Weighted Average
Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with constant learning
rate $\eta$ for the problem of prediction with expert advice is an instance of
\textsc{Mirror Descent} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
algorithm is \emph{not} scale-free.

Replacing $\frac{1}{2}\|w\|_2^2$ in equation
\eqref{equation:ada-grad-regularizer} with an arbitrary convex function $R:K
\to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{Mirror Descent}
minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero (and in particular for $t=0$). In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.

We can upper bound regret of \textsc{Scale-Free Mirror Descent} (see also \cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).  Proof is
in~\ref{section:mirror-descent-proofs}.

>>>>

\section{Scale-Free Mirror Descent}
\label{section:mirror-descent}

@@

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{MD}.  Proof
can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

@@

Several existing algorithms are special cases of \textsc{MD}.
For example, \textsc{Exponentially-Weighted Average
Forecaster}~\cite[Chapter~2]{Cesa-Bianchi-Lugosi-2006} with a constant learning
rate $\eta$ for learning with expert advice is an instance of
\textsc{MD} for the probability simplex $K = \{ w \in \R^d ~:~ w_j
\ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy regularizer $R_t(w) =
\frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $R_t:K \to \R$. Note that this
algorithm is \emph{not} scale-free.

The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
Descent} defined by the recurrence
$$
w_{t+1} = \Pi_K(w_t - \eta_t \ell_t)
$$
where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
an instance of \textsc{MD} with regularizer $R_t(w) =
\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of projected gradient descent with
step size $\eta_t = \frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized
Infinitesimal Gradient Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}.
Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
\textsc{MD} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. The diagonal version is simply the per-coordinate construction
applied to the one-dimensional version. When $d=1$, both algorithms are
equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
generalization of Zinkevich's GIGA algorithm. That is, when $\|\ell_i\|_2
= 1$ for all $i$, \textsc{MD} with regularizer
\eqref{equation:ada-grad-regularizer} coincides with GIGA.

Replacing $\frac{1}{2}\|w\|_2^2$ in equation
\eqref{equation:ada-grad-regularizer} with an arbitrary convex function $R:K
\to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free MD} with
regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{MD} minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero (and in particular for $t=0$). In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.

The theorem below upper bounds the regret of \textsc{Scale-Free Mirror Descent}
(see also
\cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).
Proof is in~\ref{section:mirror-descent-proofs}.

### 2015-12-11 preliminaries.tex
Reword and reorganize some sentences

Given $\lambda > 0$, a function $f:K \to \R$ is called \emph{$\lambda$-strongly
convex} with respect to a norm $\|\cdot\|$ if and only if, for all $x,y \in K$,
$$
f(y) \ge f(x) + \langle \grad f(x), y - x \rangle + \frac{\lambda}{2}\|x - y\|^2 \; ,
$$
where $\grad f(x)$ is any subgradient of $f$ at point $x$.

@@

\item $f^*$ is finite everywhere and differentiable.

@@

Our scale-free online learning algorithms are versions of the \textsc{Follow
The Regularized Leader} (\textsc{FTRL}) algorithm with \emph{varying
regularizers}, presented as Algorithm~\ref{algorithm:ftrl-varying-regularizer}.
The algorithm is paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex
functions $R_t:K \to \R$ called \emph{regularizers}.  Each regularizer $R_t$
can depend on the past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an
arbitrary way.  The following lemma bounds its regret.

>>>>

Given $\lambda > 0$, a function $f:K \to \R$ is called \emph{$\lambda$-strongly
convex} with respect to a norm $\|\cdot\|$ if and only if, for all $x,y \in K$,
$$
f(y) \ge f(x) + \langle \grad f(x), y - x \rangle + \frac{\lambda}{2}\|x - y\|^2 \; ,
$$
where $\grad f(x)$ is any subgradient of $f$ at the point $x$.

@@

\item $f^*$ is finite everywhere and differentiable everywhere.

@@

Our scale-free online learning algorithms are instances of \textsc{FTRL} with
\emph{varying regularizers}, presented as
Algorithm~\ref{algorithm:ftrl-varying-regularizer}.  The algorithm is
paramatrized by a sequence $\{R_t\}_{t=1}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}.  Each regularizer $R_t$ can depend on the
past loss vectors $\ell_1, \ell_2, \dots, \ell_{t-1}$ in an arbitrary way.  The
following lemma bounds its regret.
